### 2012-02-11 exam3/1.tex
nat and bool problems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}
%
Here is a simple interface and implementation for dictionaries that
associate natural numbers to strings.

\begin{verbatim}
;; A Dict is one of:
;;  - (ld-empty%)
;;  - (ld-cons% Nat String Dict)
;; implements:
;;
;; has-key? : Nat -> Boolean
;; Does this dict. have the given key?
;;
;; lookup : Nat -> String
;; Produce the value associated with the given key in this dict.
;; Assume: the key exists in this dict.
;;
;; set : Nat String -> Dict
;; Associate the given key to the given value in this dict.

(define-class ld-empty%
  (define/public (has-key? k) false)

  (define/public (set k v)
    (ld-cons% k v this)))

(define-class ld-cons%
  (fields key val rest)

  (define/public (has-key? k)
    (cond [(= (field key) k) true]
          [else ((field rest) . has-key? k)]))

  (define/public (lookup k)
    (cond [(= (field key) k) (field val)]
          [else ((field rest) . lookup k)]))

  (define/public (set k v)
    (ld-cons% k v this)))
\end{verbatim}

The dictionary data structure has some important properties, one of
which is that if we associate a value with a key and then look up that
key, we expect to get back the value we associated.  For example, for
any given dictionary, if we associate \code{1} with \code{"Fred"},
then look up \code{1}, we should get \code{"Fred"}.  More generally,
we expect this to be true for \emph{any} key-value pair, not just
\code{1} and \code{"Fred"}.

\begin{enumerate}
\item Codify the general ``set/lookup'' property, described above, as
  a predicate.  Write a successful test in terms of that predicate.
\end{enumerate}

Now consider a generalization of the ``set/lookup'' property which
says that if we set \emph{two} keys to \emph{two} values then look
each of them up, we should get back the original values.  For example, for
any given dictionary, if we associate \code{1} with \code{"Fred"} and
\code{2} with \code{"Wilma"}; then looking up \code{1} will produce
\code{"Fred"} and looking up \code{2} will produce \code{"Wilma\"}.
More generally, we can state this as a property for any two key-value
pairs.

\begin{enumerate}
\setcounter{enumi}{1}
\item Codify the general ``double-set/lookup'' property as a
  predicate.  Write a successful test in terms of that predicate.

\item Does this property hold for all dictionaries and keys-value
  pairs?  If so, why?  If not, write a counter-example, i.e.~write a
  failing test in terms of the predicate you gave in part 2.
\end{enumerate}

\ifrubric\else
\newpage
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi
\end{problem}

\begin{solution}
\begin{verbatim}
;; Part 1 
;; Dict Nat String -> Boolean            [4pt]
(define (set/lookup d k v)
  (string=? (d . set k v . lookup k) v))

;; [2pt]
(check-expect (set/lookup (ld-empty%) 1 "Fred") true)

;; Part 2
;; Dict Nat String Nat String -> Boolean   [4pt]
(define (double-set/lookup d k1 v1 k2 v2)
  (and (string=? (d . set k1 v1 . set k2 v2 . lookup k1) v1)
       (string=? (d . set k1 v1 . set k2 v2 . lookup k2) v2)))

;; [2pt]
(check-expect (double-set/lookup (ld-empty%) 1 "Fred" 2 "Wilma") true)

;; Part 3  [3pt]
;; Counter-example:
(check-expect (double-set/lookup (ld-empty%) 1 "Fred" 1 "Wilma") true)
\end{verbatim}
\end{solution}

\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}
Last night, Asumu was up late improving the {\tt class} languages.
Unfortunately, he was too tired to test his work, and his last commit
broke the numeric system.

This puts us in the dire position of no longer being able to calculate
the factorial of large numbers, and we turn to you for help.

This is the definition of a class with a {\tt fact} method:

\begin{verbatim}
;; A Fact is (new fact%)
(define-class fact%
  (define (fact n)
    (cond [(zero? n) 1]
          [else (* n (this . fact (sub1 n)))])))
\end{verbatim}

We need to rewrite this method, using the \verb|Natural| interface,
which has the following methods:

\begin{verbatim}
;; A Natural implements:
;; - zero? : -> Boolean
;; is this Natural the zero value?
;; - sub1 : -> Natural
;; produce a Natural one less than this one 
;; (or zero if this one is zero)
;; - times : Natural -> Natural
;; multiply this Natural by the given one
\end{verbatim}

Write a new definition of the \verb|fact%| class using \verb|Natural|,
and \emph{not} using the built-in numbers of the class system or
DrRacket.  You can use the constant \verb|one|, which is a \verb|Natural|
representing $1$.

\ifrubric
\else
\fi

\newpage

Now, design a data defintion that implements \verb|Natural|, and
supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.  Define the
constant \verb|one|.

It will be helpful to define the helper methods \verb|add1| and
\verb|plus|.  Also, remember that $n + m = (n-1) + m + 1$ and that $n
\cdot m = (n-1) \cdot m + n$.

\ifrubric
\else
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi
\end{problem}

\begin{solution}
\begin{verbatim}

;; Part 1: N Points
(define-class fact%
  (define (fact n)
    (cond [(n . zero?) one]
          [else (n . times (this . fact (n . sub1)))])))

;; Part 2: M points
;; A Natural is one of
;; - (new zero%)
;; - (new succ% Natural)
(define-class zero%
  (define (sub1) this)
  (define (plus n) n)
  (define (add1) (new succ% this))
  (define (times n) this)
  (define (zero?) true))

(define-class succ%
  (fields pred)
  (define (zero?) false)
  (define (add1) (new succ% this))
  (define (plus n) (this . pred . plus n . add1))
  (define (times n) (this . pred . times n . plus n))
  (define (sub1) (this . pred)))

(define one (new succ% (new zero%)))
(define zero (new zero%))

(check-expect (one . zero?) false)
(check-expect (one . sub1) zero)
(check-expect ((new succ% one) . sub1) one)
(check-expect (zero . zero?) true)
(check-expect (zero . sub1) zero)
\end{verbatim}
\newpage
\end{solution}

\begin{problem}

\textbf{Extra Credit}
\\
\noindent
After Van Horn discovered the problems with numbers, he attempted to
fix them.  Unfortunately, not only did he not fix them, he broke
booleans as well.  Fortunately, we have an idea about how you can
help, expressed in the following data definitions.

\begin{verbatim}
;; A Bool implements:
;; - branch : [Action X] [Action X] -> X
;; run the first action if this is true, otherwise run the second action

;; An [Action X] implements:
;; - run : -> X
;; do this action

\end{verbatim}

Following these interface definitions, design data and class
defintions for implementations of the \verb|Bool| interface, and
reimplement the \verb|fact%| class to use them.
You may modify your natural number implementation if neccessary for
your solution.

\ifrubric\else
\newpage
[Here is some more space for the previous problem.]
\fi

\ifrubric\else
\newpage
[Here is some more space for the previous problem.]
\fi
\end{problem}

\begin{solution}
\begin{verbatim}
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% 0) (new recur% n))))
(define-class const%
  (fields run))
(define-class recur%
  (fields n)
  (define (run) ((new fact%) . fact (this . n . sub1) . times n)))

(define-class t%
  (define (branch t e) (t . run)))
(define-class f%
  (define (branch t e) (e . run)))

(define true (new t%))
(define false (new f%))
\end{verbatim}
\end{solution}
\end{document}
